---
title: "lesson3"
author: "Sumit Kant"
date: "5 September 2016"
output: html_document
---
# Psuedo Facebook Data
THe goal here is to understand user behaviour and demographics of the facebook users. This data is similar to what data scientist at facebook use on a daily basis.

# Reading in Data
```{r}
getwd()
list.files()
setwd('D:/IIML/Crisp Analytics/Udacity/DAta analysis with R/')
```
## loading a tab separated file
```{r}

pf <- read.csv('pseudo_facebook.tsv', sep="\t")
# To check out the variable names
names(pf)
```

## Histogram of users' Birthdays

```{r}
library(ggplot2)
theme_set(theme_minimal(16)) 

qplot(x = dob_day, data=pf)
qplot(x = dob_day, data=pf) +
  scale_x_continuous( breaks = 1:31) +
  facet_wrap(~dob_month, ncol=3)
```
facet_wrap allows plotting graphs along one variable and visualize the difference across other variables
facet_grid is similar to facet_wrap as it can take two variables instead of one as an input

## Friend Count
```{r}
qplot(friend_count, data = pf) + geom_histogram()
```
This data is positively skewed and is called long tailed data.
The number of facebook friends a user has are less than 500 for most of the data.

### To limit the axes using xlim parameter
xlim uses starting point and ending point to wrap the data.

```{r}
qplot(friend_count, data = pf, xlim = c(0,1000))
```
Alternatively, we can also use scale_x_continuous to limit the data.
```{r}
qplot(friend_count, data = pf) +
  scale_x_continuous(limits = c(0,1000))
```
This allows us to make the plots in layers

## Adjusting the bin widthi
```{r}
qplot(friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,100)) 
```
## Question: Which gender group has more friends, men or women?
TO answer this we need to split the histogram into two parts
Once again we can use facet_wrap and wrap by variable 'gender'
```{r}
qplot(friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,100)) +
  facet_wrap(~gender)
```
The plot shows that the number of friends for male are higher for males than females.

***Problem***

Also plotting NA values in the gender. 

***SOlution***

Subset the data and remove NA values
```{r}
qplot(friend_count, data = subset(pf, !(is.na(gender))), binwidth = 10) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,100)) +
  facet_wrap(~gender)
```

## Statistics by 'Gender'
```{r}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```
Gives the summary statisics based on gender. The median is higher for females and the difference is 22. Also, the mean is greater for females than males

## Tenure
```{r}
qplot(x = tenure, data = pf, binwidth = 30, color = I('#1e1d1d'), fill=I('#047697'))
```
***Tenure in Years***
```{r}
qplot(x = tenure/365, data = pf, binwidth = 0.25, color = I('#1e1d1d'), fill=I('#F79420')) +
  scale_x_continuous(breaks = seq(1,7,1), limits = c(0,7)) +
  xlab('Number of years using Facebook') + 
  ylab('Number of users in sample')
```
## User ages
Histogram of users based on their age
```{r}
qplot(x = age, data = pf, binwidth = 1, color = I('#1e1d1d'), fill=I('#2ecc71')) +
  scale_x_continuous(breaks = seq(10,max(pf$age),5), limits = c(min(pf$age),max(pf$age))) +
  xlab('Age of users') + 
  ylab('Number of users in sample')
```

## Transforming data
```{r}


qplot(x=friend_count, data = pf)
summary(pf$friend_count)
summary(log10(pf$friend_count+1))

library(gridExtra)
p1 <- qplot(x = friend_count, data = pf, binwidth = 100, color = I('#1e1d1d'), fill=I('#8e44ad'))
p2 <- qplot(x = log10(friend_count+1), data = pf, binwidth = 0.05, color = I('#1e1d1d'), fill=I('#8e44ad'))
p3 <- qplot(x = sqrt(friend_count), data = pf, binwidth = 1, color = I('#1e1d1d'), fill=I('#8e44ad'))
grid.arrange(p1,p2,p3)
```
## Another way pf transforming the data

Using the ggplot function we can add layers of scale to a common plot
```{r}
q1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
q2 <- q1 + scale_x_log10()
q3 <- q1 + scale_x_sqrt()
grid.arrange(q1,q2,q3, ncol=1)

```
## Frequency Polygons
Let us compare distributions - curve drawn connecting the counts in a histogram

```{r}
qplot(friend_count, data = subset(pf, !(is.na(gender))), binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,100))

```


```{r}
qplot(x = www_likes, data = subset(pf, !(is.na(gender))),
      geom = 'freqpoly', color = gender) +
  scale_x_continuous() +
  scale_x_log10()

```
# Likes on the web
```{r}
by(pf$www_likes, pf$gender, sum)
```
Evident from the command - males likes are less than females

## BOX PLOTS
To see the difference between the distributions
```{r}
qplot(x = gender,
      y = friend_count, 
      data = subset(pf, !(is.na(gender))),
      geom = "boxplot"
      )
```

```{r}
qplot(x = gender,
      y = friend_count, 
      data = subset(pf, !(is.na(gender))),
      geom = "boxplot",
      ylim = c(0,1000))

# Another way
qplot(x = gender,
      y = friend_count, 
      data = subset(pf, !(is.na(gender))),
      geom = "boxplot") + scale_y_continuous(limits = c(0,1000))
```
Scale x continuous omits observations outside the defined limits, Thus use coord_cartesian  
```{r}
qplot(x = gender,
      y = friend_count, 
      data = subset(pf, !(is.na(gender))),
      geom = "boxplot") + coord_cartesian(ylim = c(0,1000))

```
## Who initiated more friendships?
```{r}
qplot(x = gender,
      y = friendships_initiated, 
      data = subset(pf, !(is.na(gender))),
      geom = "boxplot") + coord_cartesian(ylim = c(0,250))

by(pf$friendships_initiated, pf$gender, summary)
```



## Getting Logical
TO determine the number of users with mobile check in
```{r}
summary(pf$mobile_likes)
pf$mobile_check_in <- ifelse(pf$mobile_likes>0,1,0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
# WHat percent of Checkin using mobile
sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```


