Lesson 4
========================================================

***
```{r}
setwd("D:/IIML/Crisp Analytics/Udacity/DAta analysis with R/")
```

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv("pseudo_facebook.tsv", sep = "\t")
View(pf)
qplot(x = age, y = friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response:
* Younger Users have lots of friends
* Vertical bars that people have lied about their age like 69 and 100. These accounts tend to be of teenagers or fake accounts given the high number of friends.

***

### ggplot Syntax
Notes:
* aes - aesthatic wrapper holds x and y variables.
* Limiting the x - axes from 13-90 years since users younger than 13 cannot have fb account and over 90 years of age those are supposedly fake accounts.

```{r ggplot Syntax}
ggplot(pf, aes(age, friend_count)) + geom_point()
summary(pf$age)
#Limiting the axes
ggplot(pf, aes(age, friend_count)) + geom_point() + xlim(13,90)

```

***

### Overplotting
Notes: Overplotting means that observations are stacked over one another and it is difficult to tell the exact number of obervations. The densely populated areas is the area that is overplotted. The issue can be resolved by changing the transparecny

```{r Overplotting}
# The alpha param controls transparency.
ggplot(pf, aes(age, friend_count)) + 
  geom_point(alpha = 1/20) + 
  xlim(13,90)

# Using geom_jitter to make the age continous
ggplot(pf, aes(age, friend_count)) + 
  geom_jitter(alpha = 1/20) + 
  xlim(13,90)
```

#### What do you notice in the plot?
Response:
* Bulk of the data lies below 1000 friend count.
***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(pf, aes(age, friend_count)) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13,90) + 
  coord_trans(y = "sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(pf, aes(age, friendships_initiated)) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13,90) +
  coord_trans(y = "sqrt")
```

#### What do you notice?

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes: Analysing the variation of mean friend count over age groups makes much more sense than plotting actual data and reading from them.

```{r Conditional Means}
library(dplyr)
# some of the common functions of dplyr package are filter(), group_by(), mutate() and arrange()
# Lets group the users by age
age_groups <- group_by(pf, age)
# Summarize the data we just created and find the mean, median and number of users in each age groups
pf.fc_by_age <- summarise(age_groups, 
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
head(pf.fc_by_age)
# The data is not in any order - so arrange will help order the age groups
pf.fc_by_age <- arrange(pf.fc_by_age, age)

```
## ANOTHER WAY 
### Above code may be verbose so another way of getting the same table

```{r}
# '%.%' lets us chain functions
# pf.fc_by_age2 <- pf %>% 
#   group_by(age) %>%
#   summarise(age_groups,
#             friend_count_mean = mean(friend_count),
#             friend_count_median = median(friend_count),
#             n = n()) %>%
#   arrange(age)
```


Create your plot!

```{r Conditional Means Plot}
ggplot(pf.fc_by_age, aes(age, friend_count_mean)) + geom_line()

```

***

### Overlaying Summaries with Raw Data
Notes:  ggplot 2.0.0 changes the syntax for parameter arguments to functions when using stat = 'summary'. To denote parameters that are being set on the function specified by fun.y, use the fun.args argument, e.g.:
geom_line(stat = 'summary', fun.y = quantile,          fun.args = list(probs = .9), ... )

To zoom in, the code should use thecoord_cartesian(xlim = c(13, 90)) layer rather than xlim(13, 90) layer.

Look up documentation for coord_cartesian() and quantile() if you're unfamiliar with them. 

```{r Overlaying Summaries with Raw Data}

ggplot(aes(age, friend_count), data = pf) +
  xlim(13, 90) +
  geom_point(alpha = 0.05,
             position = position_jitter(h=0),
             color = 'green') +
  coord_trans(y = "sqrt") +
  geom_line(stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", 
            fun.y = quantile, 
            fun.args=list(probs=0.1),
            linetype = 2,
            color = "blue") +
  geom_line(stat = "summary", 
            fun.y = quantile, 
            fun.args=list(probs=0.9),
            linetype = 2,
            color = "blue") +
    geom_line(stat = "summary", 
            fun.y = median,
            color = "red")

```

## Using Coord_cartesian
Coord_cartesian acts as visual zoom on the plot rather than modifying values outside the limits. To use it we need to get rid of xlim and coord_trans layers.

```{r}

ggplot(aes(age, friend_count), data = pf) +
  coord_cartesian(xlim = c(13,70), ylim = c(0,1000)) +
  geom_point(alpha = 0.05,
             position = position_jitter(h=0),
             color = 'green') +
  geom_line(stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", 
            fun.y = quantile, 
            fun.args=list(probs=0.1),
            linetype = 2,
            color = "blue") +
  geom_line(stat = "summary", 
            fun.y = quantile, 
            fun.args=list(probs=0.9),
            linetype = 2,
            color = "blue") +
    geom_line(stat = "summary", 
            fun.y = median,
            color = "red")
ggsave('coord_cartesian_overlyaing.png', width = 12.8, height = 7.2, dpi = 300)

```


#### What are some of your observations of the plot?s
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes: Describes the strength of relationship between any two variables.

```{r Correlation}
cor.test(pf$age , pf$friend_count, method = "pearson")

# another way of calculating is using the with function
with(pf, cor.test(age, friend_count, method = "pearson"))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets

```{r Correlation on Subsets}
# The method parameter is not used here because this funtion defaluts to pearson method.
with(subset(pf, pf$age <= 70 ), cor.test(age, friend_count))

```
Notes: 
The value comes out to be  -0.17 which means age and friend_count is negatively correlated. As age increases, friend_count decreases. However, this does not imply any kind of causation.

***

### Correlation Methods
Notes: 
* "pearson", the test statistic is based on Pearson's product moment correlation coefficient cor(x, y) and follows a t distribution with length(x)-2 degrees of freedom if the samples follow independent normal distributions

* Kendall's tau or Spearman's rho statistic is used to estimate a rank-based measure of association. These tests may be used if the data do not necessarily come from a bivariate normal distribution.

***

## Create Scatterplots
Notes: Creating a scatter plot for two variables which are the number of likes recieved by the users and like recieved from desktop version of facebook as the source.

```{r}
ggplot(data = pf, aes(x = www_likes_received, y = likes_received)) + 
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = "lm", color = "red")

```

***

### Strong Correlations
Notes: High correlation does not always mean it is good. In this case one was the subset of other hence may not be very useful.

```{r Strong Correlations}
cor.test(pf$www_likes_received, pf$likes_received)
# Result: COrr = 0.948
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.


Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
# install.packages('alr3')
library(alr3)
data(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(data = Mitchell, aes(x = Month, y = Temp)) + geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(Mitchell, aes(Month, Temp)) +
  geom_point() + geom_line()
  scale_x_continuous(breaks = seq(0, 203, 12))
  
```

***

### A New Perspective
You could also get perspective on this data by overlaying each year's data on top of each other, giving a clear, generally sinusoidal graph. You can do this by using the R's modulus operator %% in your code

```{r}
ggplot(aes(x=(Month%%12),y=Temp),data=Mitchell)+ 
  geom_point() 

```

What do you notice?
Notes:


***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + geom_line()

```

### Create an age with months variable
```{r}
pf$age_with_months <- pf$age - pf$dob_month/12 + 1

View(pf)

```
***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- pf$age - pf$dob_month/12 + 1


```

Programming Assignment
```{r Programming Assignment}

age_groups <- group_by(pf, age_with_months)
pf.fc_by_age_months <- summarise(age_groups, 
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
pf.fc_by_age_months <- arrange(pf.fc_by_age_months, age_with_months)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(data = pf.fc_by_age_months, aes(x = age_with_months, y = friend_count_mean)) +
  geom_point() +
  xlim(13,71)

## ALternatively
ggplot(aes(age_with_months, friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) + geom_line()

```

***

### Smoothing Conditional Means
Notes: 
```{r Smoothing Conditional Means}
p1 <- ggplot(aes(age, friend_count_mean), data = subset(pf.fc_by_age, age < 71)) + geom_line() + geom_smooth()

p2 <- ggplot(aes(age_with_months, friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) + geom_line() + geom_smooth()

p3 <- ggplot(data  = subset(pf, age < 71), aes(x = round(age / 5)*5, y = friend_count)) + geom_line(stat = "summary", fun.y = mean)


library(gridExtra)
grid.arrange(p2, p1, p3, ncol=1)

```
p3 has less data points and wider bin widths. By doing this we will estimate the mean more precisely but potentially miss important features of the age and friend_count relationship. These three plots are the examples of bias-variance trade-off  and it is similar to the trade off we make while choosing bin width for histograms.

One way an analyst can make this trade off by using a flexible statistical model to smooth the estimates of conditional means. ggplot makes it easy to fit such models using geom_smooth. SO adding geom_smooth to the previous plots. 
***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

