---
title: "Problem Set 2 - Diamonds"
author: "Sumit Kant"
date: "25 March 2017"
output: html_document
---

Observing the dataset

```{r}
setwd("D:/IIML/Crisp Analytics/Udacity/DAta analysis with R/")
library(ggplot2) #must load the ggplot package first 
data(diamonds) #loads the diamonds data set since it comes with the ggplot package 

```

## Number of Observations
```{r}
nrow(diamonds)
```
## Number of Variables
```{r}
ncol(diamonds)
```
## Number of Ordered factors
```{r}
str(diamonds) 
```
## Best Colour proxy
```{r}
levels(diamonds$color)
# color: diamond colour, from J (worst) to D (best)
?diamonds 
```
## Create a histogram of the price of all the diamonds in the diamond data set.

```{r}
qplot(x = price, data = diamonds, binwidth = 500)
```
## Summary Statistics
```{r}
summary(diamonds$price)
```
# Diamond Counts

Diamonds costing less than 500 dollars
```{r}
nrow(diamonds[diamonds$price < 500,])
```
Diamonds costing less than 250 dollars
```{r}
nrow(diamonds[diamonds$price < 250,])
```
Diamonds costing greater than or equal to 15000 dollars
```{r}
nrow(diamonds[diamonds$price >= 15000,])
```
## Explore the largest peak in the price histogram you created earlier.

```{r}
qplot(x = price, data = diamonds, binwidth = 10) +
  scale_x_continuous(limits = c(500,1000))
```

## Facet Wrap
```{r}
qplot(x = price, data = diamonds) + facet_wrap(~cut)
```
Which cut has the highest price diamond
```{r}
by(diamonds$price, diamonds$cut, max)
```
Lowest price diamond
```{r}
by(diamonds$price, diamonds$cut, min)
```
Lowest Median price by cut
```{r}
by(diamonds$price, diamonds$cut, median)
```
## changing y-axis for different histograms

```{r}
qplot(x = price, data = diamonds) + facet_wrap(~cut, scales = "free_y")
ggsave("free_y_facetwrap_diamonds_by_Cut.png")
```
# Create a histogram of price per carat
```{r}
qplot(x = log10(price/carat), data = diamonds) + facet_wrap(~cut, scales = "free_y")
```

# Investigating the price of diamonds by color
```{r}
by(diamonds$price, diamonds$color, summary)
boxplot(price~color,data=diamonds, main="Price by color", xlab="Color", ylab="Price")
ggplot(diamonds, aes(x=color, y=price, fill=color)) + geom_boxplot()
ggsave("price_by_color.png")
```

# Investigating the price of diamonds by clarity
```{r}
by(diamonds$price, diamonds$clarity, summary)
boxplot(price~clarity,data=diamonds, main="Price by clarity", xlab="clarity", ylab="Price")
ggplot(diamonds, aes(x=clarity, y=price, fill=clarity)) + geom_boxplot()
ggsave("price_by_clarity.png")
```

## Determining IQR
```{r}
by(diamonds$price, diamonds$color, IQR)
```

# PRICE/CARAT across colors
```{r}
ggplot(diamonds, aes(x=color, y=price/carat, fill=color)) + geom_boxplot()
ggsave("price_per_carat_by_color.png")
```
# Investigating the weight of diamonds
```{r}
qplot(carat, data = diamonds, geom = "freqpoly", binwidth=0.001)
ggsave("caratcount.png")
```


