Lesson 5
========================================================

```{r}
setwd("D:/IIML/Crisp Analytics/Udacity/DAta analysis with R/Lesson 5/")
library(ggplot2)
pf <- read.csv("pseudo_facebook.tsv", sep = "\t")

```


### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(data = subset(pf, !is.na(gender)), aes(x = gender, y = age)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom= "point", shape = 4)
#shape = 4- marks an x

```
## Median Friend COunt
```{r}
ggplot(aes(age, friend_count), data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = "summary", fun.y = median)

```

## PProgramming Assignment
```{r}
library(dplyr)
age_gender_groups <- group_by(subset(pf, !is.na(gender)), age, gender)
pf.fc_by_age_gender <- summarise(age_gender_groups, 
                          mean_friend_count = mean(friend_count),
                          median_friend_count = median(friend_count),
                          n = n())
ungroup(pf.fc_by_age_gender)
arrange(pf.fc_by_age_gender, age)

```


***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(pf.fc_by_age_gender, aes(age, median_friend_count)) + 
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes: Long format includes repetition of values for same variables. They need to be converted to wwide format sometimes to get better insights.

***

### Reshaping Data
Notes: The dcast() function is used to create a data frame. For creating matrix we need to use acast() function. The left side of ~ forms the row while the right side forms the columns.

```{r}
# install.packages('reshape2')
library(reshape2)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = "median_friend_count")
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(data = pf.fc_by_age_gender.wide, aes(x = age, y = female/male)) + 
  geom_line() +
  geom_hline(yintercept = 1, color = "blue", linetype = 2)
?geom_hline
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- 2014 - ceiling(pf$tenure/365) 

head(pf$year_joined,20)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
?cut
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}

```

***

### Friending Rate
Notes:

```{r Friending Rate}

```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

